---
title: "Streamflow and fire: Case study of "
author: "Hannah"
date: "1/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages('dataRetrieval')
library(dataRetrieval) # packaage for downloading USGS stream flow data
#install.packages('lubridate')
library(lubridate) # date management
library(forecast)
library(stlplus)
library(fpp)
```

## Downloading Stream flow Data

Case study of Fawn Peek Fire Complex -- 6/30/2003
USGS gage Station Name: ANDREWS CREEK NEAR MAZAMA, WA
USGS gage ID: 12447390
*note: gage will be discontinued by the usgs Oct 2022*


```{r downloading data}
# identifying parameters for data download
siteNumber <- "11413000" # USGS gauge number
parameterCd <- "00060"  # mean daily discharge in cfs
startDate <- "1997-10-1" # ~ 5 years prefire: incorportating start of water year
                          #(this is before MODIS data):if we wanted to be in MODIS timeframe it would be 3 years pre & post
endDate <- "2008-10-1" # 5 years postfire

# download data using readNWISdv function
andrewcreek_10 <- dataRetrieval::readNWISdv(siteNumber, parameterCd, startDate, endDate)
head(andrewcreek_10) # check
colnames(andrewcreek_10)[4]<- 'discharge_cfs'
andrewcreek_10<-andrewcreek_10[,-5]

#Making time series in water year
#I am seperating the date for each column 
andrewcreek_10$year <-lubridate::year(andrewcreek_10$Date) # add year col
andrewcreek_10$month <-lubridate::month(andrewcreek_10$Date) # add month col
andrewcreek_10$day <-lubridate::day(andrewcreek_10$Date) # add day of month col
andrewcreek_10$doy <-lubridate::yday(andrewcreek_10$Date) # add day of year col
andrewcreek_10 <-dataRetrieval::addWaterYear(andrewcreek_10)
firedate<-ymd('2003-06-30')



```

## Timeseries Plots

```{r pressure, echo=FALSE}
#Plots of all years from 1998 - 2008
plot(andrewcreek_10$Date, andrewcreek_10$discharge_cfs, type='l', ylab='Discharge (cfs)', xlab='Time (days)')
abline(v=as.Date('2003-06-30'), col='red')

#Plots of just water year 2003

plot(andrewcreek_10[andrewcreek_10$Date>=as.Date('2003-10-01') & andrewcreek_10$Date<=as.Date('2004-10-01'),]$Date, andrewcreek_10[andrewcreek_10$Date>=as.Date('2003-10-01') & andrewcreek_10$Date<=as.Date('2004-10-01'),]$discharge_cfs, type='l', ylab='Discharge (cfs)', xlab='Time (days)')
abline(v=as.Date('2003-06-30'), col='red')


```
#All years of stream data

The steam gage has data from 1989-10-01 to today's date
```{r}
# identifying parameters for data download
siteNumber <- "11413000" # USGS gauge number
parameterCd <- "00060"  # mean daily discharge in cfs
startDate <- "1989-10-1" # ~ 5 years prefire: incorportating start of water year
                          #(this is before MODIS data):if we wanted to be in MODIS timeframe it would be 3 years pre & post
endDate <- "2021-01-13" # 5 years postfire

# download data using readNWISdv function
andrewcreek_all <- dataRetrieval::readNWISdv(siteNumber, parameterCd, startDate, endDate)
head(andrewcreek_all) # check
colnames(andrewcreek_all)[4]<- 'discharge_cfs'
andrewcreek_all<-andrewcreek_all[,-5]

#Making time series in water year
#I am seperating the date for each column 
andrewcreek_all$year <-lubridate::year(andrewcreek_all$Date) # add year col
andrewcreek_all$month <-lubridate::month(andrewcreek_all$Date) # add month col
andrewcreek_all$day <-lubridate::day(andrewcreek_all$Date) # add day of month col
andrewcreek_all$doy <-lubridate::yday(andrewcreek_all$Date) # add day of year col
andrewcreek_all <-dataRetrieval::addWaterYear(andrewcreek_all)

andrewcreek_all <- as.data.frame(dplyr::group_by(andrewcreek_all,waterYear) 
                      %>% mutate(dowy = seq(1:n())))

plot(andrewcreek_all$Date, andrewcreek_all$discharge_cfs, type='l', ylab='Discharge (cfs)', xlab='Time (days)')
```
##Find stream flow metrics

*Date on spring onset (Painter et al 2017)*
The onset of the rising limb is calculated here as the date on which the deviation from the running mean flow since 1 January, or date of first measurement, exceeds 1% (Figure S2). Peak flow in these rivers generally occurs sometime during the second half of May and the first half of June.

```{r}

andrewcreek2003<-andrewcreek_all[andrewcreek_all$year == 2003,] 
runningmean<-function(dataset,x){sum(dataset$discharge_cfs[1:x])/x}
andrewcreek2003$runmean<-NA
for (i in 1:length(andrewcreek2003$year)) {
  andrewcreek2003$runmean[i]<-runningmean(andrewcreek2003,i)
}
plot(andrewcreek2003$dowy,andrewcreek2003$runmean, type='l')

andrewcreek2003$permeanchange<-NA
for (i in 2:length(andrewcreek2003$year)){
andrewcreek2003$permeanchange[i]<-abs(1-andrewcreek2003$runmean[i]/andrewcreek2003$runmean[i-1])
}
plot(andrewcreek2003$dowy,andrewcreek2003$permeanchange, type='o')

```
##Creating a time series object with the ts package

```{r}
andrewcreek.ts<-ts(andrewcreek_all$discharge_cfs, start = c(1989,10), end = c(2022,01), frequency = 365)

plot(andrewcreek.ts)
andrews_add = decompose(andrewcreek.ts, type = 'additive') #The moving average window is unknown
plot(andrews_add)

#Just see one seasonal trend
plot(andrews_add$seasonal[1:365],type="l") #This seasonal trend is more gradual starting in mid-February and peaking in February
```

